Trends in atomistic simulation software usage [\versionnumber]

### 2021-10-14 manuscript.tex
revert back to bibtex/natbib
Following editorial decision, revert the bibliography format from
biber/biblatex back to bibtex/natbib in order for the format to be in
line with the other LiveCoMS articles.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% LIVECOMS ARTICLE TEMPLATE FOR BEST PRACTICES GUIDE
%%% ADAPTED FROM ELIFE ARTICLE TEMPLATE (8/10/2017)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PREAMBLE
\documentclass[9pt,review]{livecoms}
% Use the 'onehalfspacing' option for 1.5 line spacing
% Use the 'doublespacing' option for 2.0 line spacing
% Use the 'lineno' option for adding line numbers.
% Use the "ASAPversion' option following article acceptance to add the DOI and relevant dates to the document footer.
% Use the 'pubversion' option for adding the citation and publication information to the document footer, when the LiveCoMS issue is finalized.
% The 'bestpractices' option for indicates that this is a best practices guide.
% Omit the bestpractices option to remove the marking as a LiveCoMS paper.
% Please note that these options may affect formatting.

\usepackage{lipsum} % Required to insert dummy text
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\usepackage{booktabs}
\DeclareSIUnit\Molar{M}
\usepackage[italic]{mathastext}
\graphicspath{{figures/}}
%\usepackage[
%backend=biber,%
%bibstyle=numeric-comp,%
%sorting=none%
%]{biblatex}
\addbibresource{atomistic.software.bl.bib}

@@

\author[1,2,3*]{Leopold Talirz}
\author[4]{Luca M. Ghiringhelli}
\author[1,3]{Berend Smit}
%\author[1,2\authfn{1}\authfn{3}]{Firstname Middlename Familyname}
%\author[2\authfn{1}\authfn{4}]{Firstname Initials Surname}
%\author[2*]{Firstname Surname}
\affil[1]{Laboratory of Molecular Simulation (LSMO),
    Institut des Sciences et Ingenierie Chimiques,
    Valais, \'Ecole Polytechnique F\'ed\'erale de Lausanne,
    CH-1951 Sion, Switzerland}
\affil[2]{Theory and Simulation of Materials (THEOS),
    Facult\'e des Sciences et Techniques de l'Ing\'enieur,
    \'Ecole Polytechnique F\'ed\'erale de Lausanne,
    CH-1015 Lausanne, Switzerland}
\affil[3]{National Centre for Computational Design and Discovery
of Novel Materials (MARVEL), \'Ecole Polytechnique F\'ed\'erale de Lausanne,
CH-1015 Lausanne, Switzerland}
\affil[4]{The NOMAD Laboratory at the Fritz Haber Institute of the Max Planck Society and Humboldt University, Berlin, Germany}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PUBLICATION INFORMATION
%%% Fill out these parameters when available
%%% These are used when the "pubversion" option is invoked
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pubDOI{10.XXXX/YYYYYYY}
\pubvolume{<volume>}
\pubissue{<issue>}
\pubyear{<year>}
\articlenum{<number>}
\datereceived{Day Month Year}
\dateaccepted{Day Month Year}

@@

\begin{table*}
    \centering
    \begin{tabular}{llllS[table-format=7.0]} \toprule
        Software & Version & Main language& {Contributors 2020} & {Lines of code} \\ \midrule
        NWChem \cite{Apra2020}&7.0.2&Fortran& 11 &5034139\textsuperscript{*} \\
        LAMMPS \cite{Plimpton1995}&27May2021&C++& 51 & 1039872  \\
        CP2K \cite{Kuhne2020a}&8.2&Fortran & 38 & 911367 \\
        ABINIT \cite{Gonze2020}&9.4.2&Fortran& 21 & 719048 \\
Quantum ESPRESSO \cite{Giannozzi2020}&6.7.0&Fortran& 43& 604666 \\
GROMACS \cite{Abraham2015}&2021.2&C++& 25 & 589108 \\
Psi4 \cite{Smith2020}&1.3.2&C++& 34 & 496858 \\
SIESTA \cite{Garcia2020}&5.0.0-alpha&Fortran& 4 & 411829 \\
Octopus \cite{Tancogne-Dejean2020}&10.5&Fortran& 10 & 297825 \\
OpenMM \cite{Eastman2017}&7.5.1&C++& 21 & 256452 \\
xtb \cite{Bannwarth2021}&6.4.1&Fortran& 18 & 124344 \\\hline
Linux kernel&5.13&C& >4000 \cite{LinuxHistory2020} & 20904410\\
%\bottomrule
    \end{tabular}\\
    \caption{
        Counting lines of code for 11 popular open-source atomistic simulation engines (and the Linux kernel for comparison), using the latest releases as of June 2021.
        Line counts are determined by cloc v1.6.0 \cite{Danial2021} and exclude blank lines, comments, and markup languages (detailed reports in the supporting information).
        Contributors for the year 2020 were determined by counting the number of different committers to the source code from January 1st 2020 to January 1st 2021 (numbers for the Linux kernel are from 2019).
        \textsuperscript{*} Roughly 3 million lines of code of NWChem are computer-generated.
    }
    \label{tab:lines-of-code}
\end{table*}

@@

\printbibliography

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% LIVECOMS ARTICLE TEMPLATE FOR BEST PRACTICES GUIDE
%%% ADAPTED FROM ELIFE ARTICLE TEMPLATE (8/10/2017)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PREAMBLE
\documentclass[9pt,review,ASAPversion]{livecoms}
% Use the 'onehalfspacing' option for 1.5 line spacing
% Use the 'doublespacing' option for 2.0 line spacing
% Use the 'lineno' option for adding line numbers.
% Use the "ASAPversion' option following article acceptance to add the DOI and relevant dates to the document footer.
% Use the 'pubversion' option for adding the citation and publication information to the document footer, when the LiveCoMS issue is finalized.
% The 'bestpractices' option for indicates that this is a best practices guide.
% Omit the bestpractices option to remove the marking as a LiveCoMS paper.
% Please note that these options may affect formatting.

\usepackage{lipsum} % Required to insert dummy text
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\usepackage{booktabs}
\DeclareSIUnit\Molar{M}
\usepackage[italic]{mathastext}
\graphicspath{{figures/}}

@@

\author[1,2,3*]{Leopold Talirz}
\author[4]{Luca M. Ghiringhelli}
\author[1,3]{Berend Smit}
%\author[1,2\authfn{1}\authfn{3}]{Firstname Middlename Familyname}
%\author[2\authfn{1}\authfn{4}]{Firstname Initials Surname}
%\author[2*]{Firstname Surname}
\affil[1]{Laboratory of Molecular Simulation (LSMO),
    Institut des Sciences et Ingénierie Chimiques,
    Valais, \'Ecole Polytechnique F\'ed\'erale de Lausanne,
    CH-1951 Sion, Switzerland}
\affil[2]{Theory and Simulation of Materials (THEOS),
    Facult\'e des Sciences et Techniques de l'Ing\'enieur,
    \'Ecole Polytechnique F\'ed\'erale de Lausanne,
    CH-1015 Lausanne, Switzerland}
\affil[3]{National Centre for Computational Design and Discovery
of Novel Materials (MARVEL), \'Ecole Polytechnique F\'ed\'erale de Lausanne,
CH-1015 Lausanne, Switzerland}
\affil[4]{The NOMAD Laboratory at the Fritz Haber Institute of the Max Planck Society and Humboldt University, Berlin, Germany}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PUBLICATION INFORMATION
%%% Fill out these parameters when available
%%% These are used when the "pubversion" option is invoked
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pubDOI{10.XXXX/YYYYYYY}
\pubvolume{<volume>}
\pubissue{<issue>}
\pubyear{<year>}
\articlenum{<number>}
\datereceived{30 June 2021}
\dateaccepted{14 September 2021}

@@

\begin{table*}
    \centering
    \begin{tabular}{llllS[table-format=7.0]} \toprule
        Software & Version & Main language& {Contributors 2020} & {Lines of code} \\ \midrule
        NWChem \cite{Apra2020}&7.0.2&Fortran& 11 &5034139\textsuperscript{*} \\
        LAMMPS \cite{Plimpton1995}&27May2021&C++& 51 & 1039872  \\
        CP2K \cite{Kuhne2020}&8.2&Fortran & 38 & 911367 \\
        ABINIT \cite{Gonze2020}&9.4.2&Fortran& 21 & 719048 \\
Quantum ESPRESSO \cite{Giannozzi2020}&6.7.0&Fortran& 43& 604666 \\
GROMACS \cite{Abraham2015}&2021.2&C++& 25 & 589108 \\
Psi4 \cite{Smith2020}&1.3.2&C++& 34 & 496858 \\
SIESTA \cite{Garcia2020}&5.0.0-alpha&Fortran& 4 & 411829 \\
Octopus \cite{Tancogne-Dejean2020}&10.5&Fortran& 10 & 297825 \\
OpenMM \cite{Eastman2017}&7.5.1&C++& 21 & 256452 \\
xtb \cite{Bannwarth2021}&6.4.1&Fortran& 18 & 124344 \\\hline
Linux kernel&5.13&C& >4000 \cite{LinuxHistory2020} & 20904410\\
%\bottomrule
    \end{tabular}\\
    \caption{
        Counting lines of code for 11 popular open-source atomistic simulation engines (and the Linux kernel for comparison), using the latest releases as of June 2021.
        Line counts are determined by cloc v1.6.0 \cite{Danial2021} and exclude blank lines, comments, and markup languages (detailed reports in the supporting information).
        Contributors for the year 2020 were determined by counting the number of different committers to the source code from January 1st 2020 to January 1st 2021 (numbers for the Linux kernel are from 2019).
        \textsuperscript{*} Roughly 3 million lines of code of NWChem are computer-generated.
    }
    \label{tab:lines-of-code}
\end{table*}

@@

\bibliography{atomistic.software}
